---
output:
  github_document:
    html_preview: false
---

```{r include=FALSE}
knitr::opts_chunk$set(fig.path = "man/figures/", warning = FALSE)
```

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/rgexf)](https://cran.r-project.org/package=rgexf)
[![Downloads](http://cranlogs.r-pkg.org/badges/rgexf?color=brightgreen)](https://cran.r-project.org/package=rgexf)
[![Downloads](https://cranlogs.r-pkg.org/badges/grand-total/rgexf)](https://cran.r-project.org/package=rgexf)
[![R CI](https://github.com/gvegayon/rgexf/actions/workflows/ci.yml/badge.svg)](https://github.com/gvegayon/rgexf/actions/workflows/ci.yml)
[![rgexf website](https://github.com/gvegayon/rgexf/actions/workflows/website.yml/badge.svg)](https://github.com/gvegayon/rgexf/actions/workflows/website.yml)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/gvegayon/rgexf?branch=master&svg=true)](https://ci.appveyor.com/project/gvegayon/rgexf)
[![Coverage Status](https://img.shields.io/codecov/c/github/gvegayon/rgexf/master.svg)](https://codecov.io/github/gvegayon/rgexf?branch=master)
[![DOI](https://joss.theoj.org/papers/10.21105/joss.03456/status.svg)](https://doi.org/10.21105/joss.03456)

# rgexf: Build, Import and Export GEXF Graph Files <img src="man/figures/logo.svg" align="right" height="200"/>

The first R package to work with GEXF graph files (used in Gephi and
others). `rgexf` allows the user to quickly build/read graph files, including:

a. Nodes/edges attributes,

b. GEXF viz attributes (such as color, size, and position),

c. Network dynamics (for both edges and nodes, including spells), and

d. Edges weighting.

Users can build/handle graphs element-by-element or through data-frames,
visualize the graph on a web browser through ~~sigmajs javascript~~
[gexf-js](https://github.com/raphv/gexf-js) library and interact with
the igraph package.

```{r news, echo = FALSE, results='asis', }
x <- readLines("NEWS.md")
ids <- which(grepl("^#\\s*[a-zA-Z]",x))
ids <- ids[1:2] - c(0, 1)
cat(x[ids[1]:ids[2]], sep = "\n")
```

More in the [NEWS.md](NEWS.md) file.

# Installation

To install the latest version of `rgexf` you can use `devtools`

```r
library(devtools)
install_github("gvegayon/rgexf")
```

The more stable (but old) version of `rgexf` can be found on CRAN too:

```
install.packages("rgexf")
```

# Citation

```{r, comment=""}
citation(package="rgexf")
```


# Examples

## Example 1: Static net ##
```{r}
# Loading the package
library(rgexf)

# Creating a group of individuals and their relations
people <- data.frame(matrix(c(1:4, 'juan', 'pedro', 'matthew', 'carlos'),ncol=2))
people
```

```{r}
# Defining the relations structure
relations <- data.frame(matrix(c(1,4,1,2,1,3,2,3,3,4,4,2), ncol=2, byrow=T))
relations
```

```{r}
# Getting things done
write.gexf(people, relations)
```


## Example 2: Dynamic net ##

```{r}
# Defining the dynamic structure, note that there are some nodes that have NA at the end.
time<-matrix(c(10.0,13.0,2.0,2.0,12.0,rep(NA,3)), nrow=4, ncol=2)
time
```

```{r}
# Getting things done
write.gexf(people, relations, nodeDynamic=time)
```



## Example 3: More complex... Dynamic graph with attributes both for nodes and edges##

First we define dynamics

```{r}
time.nodes<-matrix(c(10.0,13.0,2.0,2.0,12.0,rep(NA,3)), nrow=4, ncol=2)
time.nodes

time.edges<-matrix(c(10.0,13.0,2.0,2.0,12.0,1,5,rep(NA,5)), nrow=6, ncol=2)
time.edges
```

Now we define the attribute values

```{r}
# Defining a data frame of attributes for nodes and edges
node.att <- data.frame(letrafavorita=letters[1:4], numbers=1:4, stringsAsFactors=F)
node.att

edge.att <- data.frame(letrafavorita=letters[1:6], numbers=1:6, stringsAsFactors=F)
edge.att

# Getting the things done
write.gexf(nodes=people, edges=relations, edgeDynamic=time.edges,
           edgesAtt=edge.att, nodeDynamic=time.nodes, nodesAtt=node.att)
```

## Example 4: Importing GEXF files

We can use the `read.gexf` function to read GEXF files into R:

```{r read-lesmiserables}
g <- system.file("gexf-graphs/lesmiserables.gexf", package="rgexf")
g <- read.gexf(g)
head(g) # Taking a look at the first handful
```

Moreover, we can use the `gexf.to.igraph()` function to convert the 
`gexf` object into an `igraph` object:

```{r igraph}
library(igraph)
ig <- gexf.to.igraph(g)

op <- par(mai = rep(0, 4)) # Making room
plot(ig)
par(op)
```

An example of the plot method using the `gexf-js` JavaScript library is available [here](https://gvegayon.github.io/rgexf/lesmiserables)

# Code of Conduct
  
We welcome contributions to `rgexf`. Whether reporting a bug, starting a
discussion by asking a question, or proposing/requesting a new feature,
please go by creating a new issue
[here](https://github.com/gvegayon/rgexf/issues) so that we can talk
about it.

Please note that the rgexf project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms


# Session info

```{r}
devtools::session_info()
```
