---
title: "rgexf: Build, Import and Export GEXF Graph Files"
output: github_document
---

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/rgexf)](https://cran.r-project.org/package=rgexf) [![Downloads](http://cranlogs.r-pkg.org/badges/rgexf?color=brightgreen)](http://cran.rstudio.com/package=rgexf)
[![Downloads](https://cranlogs.r-pkg.org/badges/grand-total/rgexf)](http://cran.rstudio.com/package=rgexf)
[![Travis-CI Build Status](https://travis-ci.org/gvegayon/rgexf.svg?branch=master)](https://travis-ci.org/gvegayon/rgexf)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/gvegayon/rgexf?branch=master&svg=true)](https://ci.appveyor.com/project/gvegayon/rgexf)
[![Coverage Status](https://img.shields.io/codecov/c/github/gvegayon/rgexf/master.svg)](https://codecov.io/github/gvegayon/rgexf?branch=master)
[![lifecycle](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://www.tidyverse.org/lifecycle/#stable)


The first R package to work with GEXF graph files (used in Gephi and others). Using XML library, it allows the user to easily build/read graph files including attributes, GEXF viz attributes (such as color, size, and position), network dynamics (for both edges and nodes, including spells) and edges weighting. Users can build/handle graphs element-by-element or massively through data-frames, visualize the graph on a web browser through ~~sigmajs javascript~~ [gexf-js](https://github.com/raphv/gexf-js) library and interact with the igraph package.

## News ##

- [2020-02-11] Getting ready to put the package on CRAN again (was taken out in the last version).
- [2016-11-08] Restarting the project.
- [2015-02-03] Version 0.15.2.3 of rgexf is on CRAN. Just updating emails...
- [2014-03-10] Version 0.14.9 of rgexf is on CRAN! solves issues 15-18. Thanks =).
- [2013-08-07] Version 0.13.8 of rgexf is on CRAN! New plot.gexf method and igraph
  integration working =).
- [2013-05-09] Version 0.13.05 of rgexf (transitory) solves issues N 9 & 10. Looking
  forward for the next CRAN version.
- [2013-03-14] Version 0.13.03 of rgexf is on its way to CRAN. It now supports
  working with spells!
- [2013-01-04] Version 0.13.01 of rgexf is on its way to CRAN. Significant
  improvements and new features!
- [2012-06-19] Version 0.12.06 of rgexf is on CRAN! Now it can be directly download
  from R.
- [2012-03-29] Version 0.12.03.29 of rgexf has been released including many bug
  fixes. Please download the latest version to check it out.

## Installation

To install the latest version of `rgexf` you can use `devtools`

```r
library(devtools)
install_github("gvegayon/rgexf")
```

The more stable (but old) version of `rgexf` can be found on CRAN too:

```
install.packages("rgexf")
```

## Example 1: Static net ##
```{r}
# Loading the package
library(rgexf)

# Creating a group of individuals and their relations
people <- data.frame(matrix(c(1:4, 'juan', 'pedro', 'matthew', 'carlos'),ncol=2))
people
```

```{r}
# Defining the relations structure
relations <- data.frame(matrix(c(1,4,1,2,1,3,2,3,3,4,4,2), ncol=2, byrow=T))
relations
```

```{r}
# Getting things done
write.gexf(people, relations)
```


## Example 2: Dynamic net ##

```{r}
# Defining the dynamic structure, note that there are some nodes that have NA at the end.
time<-matrix(c(10.0,13.0,2.0,2.0,12.0,rep(NA,3)), nrow=4, ncol=2)
time
```

```{r}
# Getting things done
write.gexf(people, relations, nodeDynamic=time)
```



## Example 3: More complex... Dynamic graph with attributes both for nodes and edges##

First we define dynamics

```{r}
time.nodes<-matrix(c(10.0,13.0,2.0,2.0,12.0,rep(NA,3)), nrow=4, ncol=2)
time.nodes

time.edges<-matrix(c(10.0,13.0,2.0,2.0,12.0,1,5,rep(NA,5)), nrow=6, ncol=2)
time.edges
```

Now we define the attribute values

```{r}
# Defining a data frame of attributes for nodes and edges
node.att <- data.frame(letrafavorita=letters[1:4], numbers=1:4, stringsAsFactors=F)
node.att

edge.att <- data.frame(letrafavorita=letters[1:6], numbers=1:6, stringsAsFactors=F)
edge.att

# Getting the things done
write.gexf(nodes=people, edges=relations, edgeDynamic=time.edges,
           edgesAtt=edge.att, nodeDynamic=time.nodes, nodesAtt=node.att)
```

# Session info

```{r}
devtools::session_info()
```


