---
title: "Visualizing GEXF graphs with sigma.js"
author: "George G. Vega Yon"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Visualizing GEXF graphs with sigma.js}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(comment = "#>")
```

# Overview

`rgexf` ships with an interactive graph viewer built on top of
[sigma.js v3](https://www.sigmajs.org/) and
[graphology](https://graphology.github.io/).  The widget is zero-dependency
from the user's perspective: the required JavaScript libraries are bundled
inside the package, so no internet connection is needed at runtime.

The main entry point is `plot()` — the standard R S3 method dispatches to
the sigma.js widget automatically for `gexf` objects.  The underlying
`sigmajs()` function is also exported for cases where you need extra
arguments such as `borderColor` or explicit `width`/`height`.

---

# 1  A simple coloured graph

## 1.1  Creating the graph

We start by building a small three-node graph from scratch.  `rgexf` stores
visual properties (colour, position, size) through the `nodesVizAtt` argument
of `gexf()`.

```{r simple-graph}
library(rgexf)

# Node and edge tables
nodes <- data.frame(
  id    = 1:3,
  label = c("Alice", "Bob", "Carol")
)
edges <- data.frame(
  source = c(1L, 2L, 1L),
  target = c(2L, 3L, 3L)
)

# Visual attributes
node_colors <- data.frame(
  r = c(220L,  66L,  40L),
  g = c( 50L, 133L, 167L),
  b = c( 47L, 244L, 240L),
  a = c(  1,    1,    1)
)
node_positions <- data.frame(
  x = c(-1,  1,  0),
  y = c( 0,  0,  1.5),
  z = c( 0,  0,  0)
)
node_sizes <- c(8, 8, 8)

g <- gexf(
  nodes        = nodes,
  edges        = edges,
  nodesVizAtt  = list(
    color    = node_colors,
    position = node_positions,
    size     = node_sizes
  )
)
```

## 1.2  Plotting the graph

Calling `plot()` on a `gexf` object renders the sigma.js widget directly.

```{r simple-plot}
plot(g)
```

By default sigma.js renders nodes as plain filled circles — the node's
`color` fills the entire disc.  You can add an optional border ring by
passing `borderColor` (any extra arguments are forwarded to `sigmajs()`):

```{r simple-plot-border, eval=FALSE}
# White border ring (15 % of the node radius)
plot(g, borderColor = "#ffffff", borderSize = 0.15)

# Dark charcoal border, slightly thicker
plot(g, borderColor = "#333333", borderSize = 0.2)
```

---

# 2  The Les Misérables network

The package includes a classic co-occurrence network from Victor Hugo's
*Les Misérables*, originally prepared for Gephi.  It already contains
`viz:color`, `viz:position`, and `viz:size` attributes, so the widget picks
them up automatically.

## 2.1  Reading the file

```{r read-lesmi}
lesmi_path <- system.file("gexf-graphs/lesmiserables.gexf", package = "rgexf")
lesmi      <- read.gexf(lesmi_path)
lesmi
```

## 2.2  Default plot

```{r lesmi-plot-default}
plot(lesmi)
```

The colours come directly from the GEXF file.  Each character group is
assigned a distinct hue by the original dataset, so the community structure
is immediately visible.

## 2.3  Adding node borders for contrast

The default rendering is sigma.js's plain filled circle.  If you want to
separate overlapping nodes or simply add visual polish, pass a `borderColor`.
A white outer ring works well on light backgrounds:

```{r lesmi-plot-border}
plot(lesmi, borderColor = "#ffffff", borderSize = 0.15)
```

The `borderColor` argument accepts any CSS colour string (`"#rrggbb"`,
`"rgb(r,g,b)"`, named colours such as `"white"`, etc.).  The `borderSize`
argument is a number between 0 and 1 representing the border width as a
fraction of the node's radius — `0.15` means the ring takes up 15 % of the
node radius.

For a dark-themed background, try a darker value:

```{r lesmi-plot-dark, eval=TRUE}
plot(lesmi, borderColor = "#222222", borderSize = 0.1)
```

---

# 3  Using the legacy gexf-js viewer

The older `gexf-js`-based viewer is preserved as `plot_gexfjs()` for users
who need the file-server workflow:

```{r gexfjs-legacy, eval=TRUE}
plot_gexfjs(lesmi)
```

This starts a local HTTP server and opens the graph in the default browser,
which is useful for standalone HTML reports or for users who prefer the
original gexf-js interface. You can also embed the graph in a Shiny app or R Markdown using the `gexfjs()` function:

```{r gexfjs-embed, eval=TRUE}
gexfjs(lesmi)
```

---

# Session info

```{r session-info}
sessionInfo()
```
